### 2015-03-26 draft.tex
Initial draft.

%===============================================================================
% $Id: ifacconf.tex 19 2011-10-27 09:32:13Z jpuente $  
% Template for IFAC meeting papers
% Copyright (c) 2007-2008 International Federation of Automatic Control
%===============================================================================
\documentclass{ifacconf}

\usepackage{graphicx}      % include this line if your document contains figures
\usepackage{natbib}        % required for bibliography
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}

\usepackage{todonotes}
\newcommand{\NOTE}[1]{\todo[inline]{#1}}

\newcommand*{\Resize}[1]{\resizebox{\columnwidth}{!}{$#1$}}

\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\providecommand{\norm}[1]{\left\|#1\right\|}
\providecommand{\abs}[1]{\left|#1\right|}
\providecommand{\conv}{\text{conv}}

\usepackage{xcolor}
\def\edit{\textcolor{blue}}
\allowdisplaybreaks[4]

%===============================================================================
\begin{document}
\begin{frontmatter}

% predefined environments
%\begin{thm} ... \end{thm}		% Theorem
%\begin{lem} ... \end{lem}		% Lemma
%\begin{claim} ... \end{claim}	% Claim
%\begin{conj} ... \end{conj}	% Conjecture
%\begin{cor} ... \end{cor}		% Corollary
%\begin{fact} ... \end{fact}	% Fact
%\begin{hypo} ... \end{hypo}	% Hypothesis
%\begin{prop} ... \end{prop}	% Proposition
%\begin{crit} ... \end{crit}	% Criterion
%\begin{rem} ... \end{rem}    % Remark
%\begin{pf} ... \end{pf}      % Proof
%\begin{ack} ... end{ack}     % Acknowledgement

\title{Robust receding horizon control for linear systems with state dependent disturbance}

\thanks[footnoteinfo]{Supported by the Martin Senior Scholarship of Worcester College, Oxford.}

\author[Oxford]{Rainer M. Schaich\thanksref{footnoteinfo}} 
\author[Oxford]{Mark Cannon}

\address[Oxford]{Department of Engineering Science, University of Oxford, OX1 3PJ, UK. (e-mail: 
rainer.schaich@eng.ox.ac.uk, mark.cannon@eng.ox.ac.uk).}

\begin{abstract}                % Abstract of not more than 250 words.

\end{abstract}

\begin{keyword}
-
\end{keyword}

\end{frontmatter}
%===============================================================================

\section{Introduction}\label{sec:intro}
Robust model predictive control (RMPC) algorithms compute a control law optimising a performance
criterion subject to state and control constraints for a perturbed system, for all possible realisations of the
perturbation. \cite{Witsenhausen:1968} proposes to formulate the RMPC problem as a game in which the adversary 
tries to maximise the effect of the uncertainty on the cost and the controller chooses its strategy to minimise 
the cost for the worst case system, this formulation is widely accepted, see e.g.~\cite{Bemporad:1999,Bemporad:2003,Lee:1997}.
The worst case system can either be considered to be open-loop, i.e. the controller chooses its strategy for the 
whole horizon before the adversary decides its strategy, or closed-loop, i.e. at each stage along the horizon 
the controller and the adversary play the game for only that stage given the outcome of the previous stage.
Theoretical properties of these controllers are well understood, see e.g.~\cite{Mayne:2000}. However, the
numerical computation of robust model predictive controllers still poses a challenge and only few tractable 
solutions have been proposed.

Numerous strategies have been proposed to solve RMPC problems for linear systems subject to linear state and
input constraints and linearly constrained additive uncertainty.
\NOTE{Do we need to constrain ourselves to additive here? In the literature there are algorithms that cope with
multiplicative just as well..?!}
\cite{Scokaert:1998} propose a scenario-based approach to solve the sequence of min-max problems, however the
complexity grows exponentially with the length of the horizon. A parametrised algorithm was proposed by~\cite{Rakovic:2012}
which can be used to reduce the RMPC problem to a single linear program that can be solved online, bit this 
parametrisation is in general suboptimal. Explicit RMPC algorithms solve the min-max problems offline and exploit 
the structure of the solution using results from multi-parametric programming to parametrise the optimal feedback as 
piecewise affine functions of the state, see e.g.~\cite{Bemporad:2003,Diehl:2004}. However, this requires extensive offline
computation and becomes inefficient for a high dimensional state space. Another approach exploiting the structure
of the solution in the same way but with less redundancy was proposed by~\cite{Buerger:2011} and extended in~\cite{Buerger:ACC},
where an active set solver is designed to solve the sequence of min-max problems explicitly online. In all 
aforementioned papers and, to the authors' best knowledge, in published literature on algorithmic solutions to the RMPC 
problem, the disturbance sets are assumed to be constant, i.e. independent of the system state and control inputs.
Recent algorithmic advances allow us to treat polytopic disturbance sets which depend in an affine way on the system state
and the control input. \cite{Schaich:2015} provide an algorithmic and analytical framework to
embed affine state and input dependent disturbance constraints into the framework of the active set RMPC solver
framework as proposed by~\cite{Buerger:2011}. Disturbance constraints varying with the state and input can be used
to model multiplicative uncertainty as well as linearisation errors which will be shown in an example.

After describing the considered control problem in section~\ref{sec:problem:statement} we summarise preliminary 
results which are necessary to formulate the considered problem explicitly (section~\ref{sec:preliminaries}).
Section~\ref{sec:gen:mpQP} discusses the recursive solution for equality constrained min-max recursions, one of 
the essential parts of the active set solver, it is applied to the considered min-max problem in section~\ref{sec:applied:mpQP}
assuming the active set of constraints is known. A line search to determine the active set of constraints is presented
in section~\ref{sec:line:search}, the application of the proposed algorithm to an example system is presented in 
section~\ref{sec:example}. Conclusions are given in section~\ref{sec:conclusion}.

\section{Problem Statement}\label{sec:problem:statement}
The set of systems to which a min-max robust receding horizon control scheme can be applied online 
without approximating the solution is bounded by the system complexity. The way to solve min-max problems we 
adopt here, is to use the structure of the solution and derive multi-parametric programs that can be solved 
efficiently. In order to be able to obtain a closed form solution of the min-max program simple optimisation 
formulations have to be available, in particular we consider linear constraints. In presence of state constraints 
the algorithmic treatment of state constraints in min-max problems stagnated after it was introduced for fixed 
polytopic disturbance sets by~\cite{Bertsekas:1971}. Recently~\cite{Schaich:2015} proposed an algorithm to compute the
so called maximal robust positive invariant (MRPI) sets (see~\cite{blanchini:2007}) for linear systems subject to state 
and input dependent disturbance constraints is presented. In this paper we develop the algorithm even further allowing 
us to compute polytopic constraints necessary to satisfy state constraints robustly in this setup. We present a 
robust receding horizon control scheme for
%
\begin{equation}\label{eq:system:equation}
	x^+ = A x + B u + D w,
\end{equation}
%
where $A,B,D$ are real matrices of appropriate dimension, $x, x^+\in\mathcal X\subseteq\mathbb R^n$ denote the state 
and successor state respectively, $u\in\mathcal U=\{u:Fu\leq{\bf{1}}\}\subset\mathbb R^p$ the input and $w\in\mathcal W(x,u)
\subset\mathbb R^q$ the disturbance with the state and input dependent disturbance set $\mathcal W(x,u)$. The disturbance 
set $\mathcal W(x,u)$ is defined by
%
\begin{equation}\label{eq:PWA:def:disturbance}
	\mathcal W(x,u) = \{w: Gw \leq H(x,u)\},
\end{equation}
%
where $H(x,u)$ is an element-wise convex piecewise affine function. The min-max optimal control problem
we will optimise is given by the $N$-stage min-max program
%
\begin{subequations}\label{eq:the:problem}%
\begin{equation}
	J_{m}^\ast(x) = \min_u \max_{w,x^+} \tfrac{1}{2}\bigl(\norm{x}_Q^2+\norm{u}_R^2-\gamma^2
    \norm{w}^2\bigr) + J_{m-1}^\ast(x^+)
\end{equation}
subject to
\begin{gather}
	x^+ = A x + B u + D w\\
	u\in\mathcal U \\
	w\in\mathcal W(x,u) \\
	(x,u)\in\mathcal Z_m
\end{gather}
\end{subequations}
%
where we enforce the stage wise state and input constraints $\mathcal Z_m$ in such a way that the terminal state,
i.e. $x^+$ at $m=1$, is contained in the MRPI set $x^+\in\mathcal X^\infty$ for the closed-loop 
system governed by a fixed state feedback controller $u=Kx$. The MRPI set $\mathcal X^\infty$ for this considered
scenario is therefore given by
%
\begin{equation}\label{eq:definition:MRPI:set}
	\mathcal X^\infty = \{x: (A+BK)x + Dw \in \mathcal X^\infty \forall w\in\mathcal W(x,Kx)\}.
\end{equation}
%
subsequently the stage wise constraints $\mathcal Z_m$ are defined recursively as
%
\begin{subequations}\label{eq:definition:stage:constraints}
\begin{align}
	\mathcal Z_m &= \{(x,u)\,: u\in\mathcal U \nonumber\\
	&\quad\ \wedge \; A x + Bu + Dw\in\mathcal X_{m-1}\; \forall\, w\in\mathcal W(x,u)\} 
	\label{eq:definition:mixed:stage:constraint}\\
	\mathcal X_m &= \{x : \exists\, u\in\mathcal U, (x,u)\in\mathcal Z_m\}
	\label{eq:definition:state:stage:constraint}\\
	\mathcal X_0 &= \mathcal X^\infty
\end{align}
\end{subequations}

\section{Preliminaries}\label{sec:preliminaries}
In this section we summarise preliminaries necessary to formulate the considered problem~\eqref{eq:the:problem}
in a convenient way. The point-wise polytopic set defined by~\eqref{eq:PWA:def:disturbance} with a elementwise convex 
piece-wise affine right-hand-side $H$ can be shown to lead to a polytopic MRPI set, see~\cite{Schaich:2015} for an 
elaborate discussion. We summarise the result as we require it in
%
\begin{lem}\label{thm:MRPI:set:is:polytopic}
Let $\mathcal X \subseteq \{x:\Gamma x\leq{\bf{1}}\wedge-\Gamma x\leq{\bf{1}}\}$ polyhedral, and the pair $(A+BK,\Gamma)$ be
observable, furthermore let $\mathcal W(x,Kx)$ be defined as in~\eqref{eq:PWA:def:disturbance} with a elementwise convex piecewise
affine function $H(x,Kx)$ which is bounded for all finite $x$. Then the MRPI set $\mathcal X^\infty$ is a polytope, i.e.
it has the representation 
\begin{equation}\label{eq:halfspace:representation:MRPI}
	\mathcal X^\infty=\{x:\Lambda_i x\leq \lambda_i, \forall\, i\in\mathcal I^\infty\}.
\end{equation}
\end{lem}
%
We need to emphasise the following 
\begin{rem}\label{rem:vertex:representation}
Notice that one essential assumption in Lemma~\ref{thm:MRPI:set:is:polytopic} is that $\mathcal W(x,u)$ is a polytope
for all finite $(x,u)$, i.e. it has a convex hull representation $\mathcal W(x,u) = \conv_k\{w_k(x,u)\}, k\in\{1,\dots,r\}$ where each 
vertex~$w_k(x,u)$ is a piecewise affine function in $(x,u)$,i.e.
\begin{equation}
\Resize{
	w_k(x,u) = \max\{W_{k,1}^x x + W_{k,1}^u u + w_{k,1},W_{k,2}^x x + W_{k,2}^u u + w_{k,2},\dots\}
}
\end{equation}
where the maximisation is done elementwise.
\end{rem}
%
We now derive the representation of $\mathcal Z_m$ and $\mathcal X_m$ as defined in~\eqref{eq:definition:stage:constraints}.
According to Lemma~\ref{thm:MRPI:set:is:polytopic} the MRPI set $\mathcal X^\infty$ is polytopic and has the 
representation~\eqref{eq:halfspace:representation:MRPI}, the recursion~\eqref{eq:definition:mixed:stage:constraint}
therefore defines $\mathcal Z_1$ as
\begin{equation}\label{eq:derivation:of:mpLP:stage:constraints}
\Resize{
\begin{aligned}
	\mathcal Z_1 =& (\mathbb R^n\times\mathcal U)\cap\\
	&\{(x,u):\Lambda_{0,i}(Ax+Bu+Dw)\leq\lambda_{0,i} \forall w\in\mathcal W(x,u),i\in\mathcal I_0\}\\
	=&(\mathbb R^n\times\mathcal U)\cap\\
	&\left\{\vphantom{\max_{w}}\right.
	(x,u):\max_{w\in\mathcal W(x,u)}\Lambda_{0,i}(Ax+Bu+Dw)\leq\lambda_{0,i}\forall i\in\mathcal I_0
	\left.\vphantom{\max_w}\right\}\\
	=&(\mathbb R^n\times\mathcal U)\cap\\
	&\left\{\vphantom{\max_{w}}\right.
	\Lambda_{0,i}(Ax+Bu)+ \underbrace{\max_{j\in\{1,\dots,r\}}\Lambda_{0,i} Dw_j(x,u)}_{(I)}\leq\lambda_i\forall i\in\mathcal I_0
	\left.\vphantom{\max_w}\right\}
\end{aligned}
}
\end{equation}
The maximisation~$(I)$ in~\eqref{eq:derivation:of:mpLP:stage:constraints} is a multi-parametric
linear program, its solution is piecewise affine. Therefore the set $\mathcal Z_1$ has a representation
$\mathcal Z_1=\{(x,u):\Xi^x_{1,i}x+\Xi^u_{1,i}u\leq\xi_{1,i},i\in\mathcal J_1\}$, its projection~$\mathcal X_1$ 
onto~$\mathbb R^n$ is therefore has the representation~$\mathcal X_1 = \{x:\Lambda_{1,i} x\leq\lambda_{1,i},i\in\mathcal I_1\}$.
From here the determination of $\mathcal Z_k$ and $\mathcal X_k$ follows recursively. We are now able to formulate 
all constraints in~\eqref{eq:the:problem} as inequalities. In the next section we discuss a method of solving
min-max programs such as~\eqref{eq:the:problem} efficiently.

\section{Solving equality constrained multi-parametric quadratic programs}\label{sec:gen:mpQP}
In this section we discuss equality constraint multi-parametric quadratic programs (mpQP) 
(see~\cite{bemporad02,Tondel:2003} for an extended discussion), as they are essential in the treatment 
of the considered min-max program using an active set solver. Note that the notation in this section does 
not refer to the notation in the preceding and proceeding sections.

Consider the primal equality-constrained concave mpQP
\begin{equation}\label{mpQP:primal}
	V(\theta) = \left\{\begin{split}
	\max_x & \;\frac{1}{2} x^T W x + c^Tx + d\\
	\text{s.t.} &\; Hx  = f +S\theta
	\end{split}\right.
\end{equation}
for $W\preceq 0.$ The Lagrangian function associated with~\eqref{mpQP:primal} is
\begin{equation}
	L(x,\lambda,\theta)  = \frac{1}{2} x^T W x + c^Tx + d + \lambda^T(Hx - f - S\theta)
\end{equation}
and the dual problem is the unconstrained mpQP
\begin{equation}\label{mpQP:dual}\begin{split}
	\min_\lambda &\; \mathcal D(\lambda,\theta)= \min_\lambda-\frac{1}{2}\lambda^T H W^{-1}H^T \lambda 
	-(HW^{-1}c\\ &+ f + S\theta)^T\lambda
	-\frac{1}{2}c^TW^{-1}c +d
\end{split}\end{equation}
To solve~\eqref{mpQP:primal} we consider the first order optimality conditions, see e.g.~\cite{Fletcher:2000}:
\begin{equation}\label{kkt:conditions}
	\begin{aligned}
        W x + H^T\lambda & = -c \\
        H x &= f + S\theta.
	\end{aligned}
\end{equation}

\begin{thm}\label{thm:mpQP:solution}
The solution of~\eqref{kkt:conditions} has the representation
\begin{equation}\label{mpQP:solution}
	\begin{split}
		x &= M_\theta \theta + m\\
		\lambda &= N_\theta \theta + N_\beta \beta + n
	\end{split}
\end{equation}
where $N_\beta\neq 0$ only if $H$ is rank deficient, in which case the rows
of $N_\beta$ form a basis of the kernel of $H$, i.e. $H^TN_\beta=0$
\end{thm}
%
The matrices~$M_\theta$, $N_\theta$, $N_\beta$ as well as the vectors $m$ and $n$
are constant, the variable $\beta$ is used to parametrise the solution in the null space of~\eqref{kkt:conditions}.
%
\begin{pf}
There are two cases of rank deficient $H$: rank$(H)=n_x<n_\lambda$ and
rank$(H)<\min\{n_x,n_\lambda\}$.
Using the QR-decomposition (see e.g.~\cite{Nocedal:2006}) of $H=(Q_1,Q_2)\cdot(R^T,0)^T$ the solution 
is given by $(m^T,n^T)^T = \Sigma (-c^T,f^T)^T$, $(M_\theta^T,N_\theta^T)^T = 
\Sigma (0,S^T)^T$ and $N_\beta=Q_2$, where 
\begin{equation}
	\Sigma = \left(\begin{array}{cc}
	0 & R^{-1}Q_1^T \\
	Q_1 R^{-T} & -Q_1R^{-T}WR^{-1}Q_1^T
	\end{array}\right)
\end{equation}
in the case where the number of equality constraints is larger than the dimension of
the primal variable, i.e. rank$(H)=n_x$. If rank$(H)<\min\{n_x,n_\lambda\}$, so that $R$ is
not invertible, then
\begin{equation}
	\Sigma = \left(\begin{array}{cc}
	W^{-1}-W^{-1}R^T\Delta RW^{-1} & W^{-1}R^{T}\Delta Q_1^{T}\\
	Q_1\Delta R W^{-1} & -Q_1\Delta Q_1^T
	\end{array}\right)
\end{equation}
with $\Delta = (RW^{-1}R^T)^{-1}$. This concludes the proof.
\end{pf}
%
Since the degeneracy variable $\beta$ does not appear in the
primal optimiser, it does not affect the primal objective value of~\eqref{mpQP:primal}.
However, substituting~\eqref{mpQP:solution} into the dual objective~\eqref{mpQP:dual} yields
\begin{multline}
    \mathcal D(\lambda(\theta,\beta),\theta) =  -\frac{1}{2}(N_\theta \theta + N_\beta \beta + n)^T H W^{-1}H^T 
	(N_\theta \theta\\ + N_\beta \beta + n)
	-(AW^{-1}c + 
	f + S\theta)^T(N_\theta \theta + N_\beta \beta + n)\\ -\frac{1}{2}c^TW^{-1}c,
\end{multline}
which is unbounded if
\begin{equation}
	(f^T+\theta^TS^T)N_\beta \neq 0
\end{equation}
and can be made arbitrarily small by the choice of $\beta$. Therefore we restrict $\theta$ to satisfy 
\begin{equation}\label{mpQP:compatability:assumption}
	N_\beta^Tf = -N_\beta^TS\theta,
\end{equation} 
so that the dual mpQP has a unique solution
and the choice of $\beta$ does not affect either the primal nor the dual cost and can be used for
any purpose. For a general mpQP~\eqref{mpQP:compatability:assumption} is a restriction on the set
of parameters for which~\eqref{mpQP:primal} has a solution of the form~\eqref{mpQP:solution}.
In our multi-stage min-max setup, the parameter is the optimiser of a subsequent optimisation
problem and can be ensured to satisfy~\eqref{mpQP:compatability:assumption} by the so called 
compatibility constraint. The optimal objective of~\eqref{mpQP:primal} is given by
\begin{equation}\label{mpQP:cost:to:go}\begin{split}
	V(\theta) = &\frac{1}{2}\theta^T M_\theta^T W M_\theta \theta + c^TM_\theta \theta
\\&+ m^TWM_\theta+\frac{1}{2}m^T Wm + c^Tm\\
=&\frac{1}{2} \theta^T W_\theta \theta + c^T_\theta \theta + d_\theta.
\end{split}\end{equation}
%
Assume now that we want to solve
\begin{equation}\label{mpQP:second:stage}
	J(\phi) = \left\{\begin{array}{rl}
	\min_\theta & \frac{1}{2}\theta^T\Phi\theta + g^T\theta + f + V(\theta)\\
	\text{s.t.} & C\theta = e+T\phi
	\end{array}\right.
\end{equation}
where the equality constraints include~\eqref{mpQP:compatability:assumption}. Since we know
that the optimal cost-to-go $V(\theta)$ has the structure~\eqref{mpQP:cost:to:go}
the problem~\eqref{mpQP:second:stage} reduces to a problem like~\eqref{mpQP:primal},
which yields a solution like~\eqref{mpQP:solution}, i.e.
\begin{equation}
\begin{split}
	\theta &= K_\phi \phi + k\\
	\eta &= L_\phi \phi + L_{\hat\beta} \hat\beta + l
\end{split}
\end{equation}
with the dual variable~$\eta$, as long as $\Phi+W_\theta\succ 0$. Finally we note 
that~\eqref{mpQP:compatability:assumption} is only nontrivial if~$H$ is rank deficient and only 
introduces as many constraints on~$\theta$ as there are linearly
dependent rows in~$H$.

In the solution of recursive min-max programs such as~\eqref{eq:the:problem} we use the composition of
the stage optimisers, i.e. we use $x(\theta(\phi)) = M_\theta K_\phi \phi + (M_\theta k + m)$.
Notice that this composition of optimisers is again an affine function of the parameter, in particular
we will exploit the fact that changing the parameter along a line will translate to a line in the optimiser
space as long as no activations or deactivations in the constraints occur. We will revisit this fact
when we perform a line search in order to update the active constraints.

\section{Solving min-max quadratic programs as multi-parametric quadratic programs}\label{sec:applied:mpQP}
In this section we apply the results derived in section~\ref{sec:gen:mpQP} to
the considered min-max robust control problem~\eqref{eq:the:problem}. We 
define a recursion and assume the active set is known to obtain a closed solution for~\eqref{eq:the:problem}.
%
To use an active set solver for the problem:
\begin{subequations}\label{eq:the:problem:reformulated}%
\begin{equation}
	J_{m}^\ast(x) = \min_u \max_{w,x^+} \tfrac{1}{2}\bigl(\norm{x}_Q^2+\norm{u}_R^2-\gamma^2
    \norm{w}^2\bigr) + J_{m-1}^\ast(x^+)
\end{equation}
subject to
\begin{gather}
	x^+ = A x + B u + D w
\\
\label{eq:dist:cons}
	G w \leq  H(x,u)
\\
\label{eq:state:constraints}
	\Xi^x_m x + \Xi^u_m u \leq \xi_m
\\
	\hat C_m x^+ = - {\bf{1}}
\\
	C_m^x x + C_m^u u  = {\bf{1}}
\end{gather}
\end{subequations}
with~$J_0^\ast(x^+) = \frac{1}{2}\norm{x^+}_P^2$ and $m=1,\dots,N$. Notice that explicitly using
$u\in\mathcal U$ is no longer necessary as satisfaction of the input constraints is intrinsic
if $(\cdot,u)\in\mathcal Z_m$.
We split the sequence of  min-max problems into a sequence of multi-parametric 
quadratic programs. Starting with the maximisation at stage $m=1$ and using a backward sweep approach (see e.g.~\cite{Bryson:1975}), 
we obtain
\begin{equation}
	\hat J_m^\ast(x,u) = \left\{\begin{split}
    \max_{w,x^+} & -\frac{\gamma^2}{2}\norm{w}^2 + J_{m-1}^\ast(x^+)\\
    \text{s.t.} & \begin{array}{rcl}
    Gw & \leq & H(x,u)\\
    x^+ & = & Ax + Bu + Dw\\
    \hat C_m x^+ &=& -{\bf{1}}
    \end{array}
    \end{split}\right.
\end{equation}
and
\begin{equation}
	J_m^\ast(x) = \left\{\begin{split}
    \min_{u} & \frac{1}{2}\left(\norm{x}^2_Q + \norm{u}_R^2\right) + \hat J_m^\ast(x,u)\\
    \text{s.t.} & \begin{array}{ccccl}
    \Xi^x_m x &+& \Xi^u_m u &\leq &\xi_m\\
    C_m^x x &+ & C^u_m u & = & {\bf{1}}
    \end{array}
    \end{split}
    \right.
\end{equation}
Suppose the active set is given, i.e.~it is known which inequality constraints hold with equality. 
For notational convenience we denote active inequalities with $(\cdot)^\prime$, e.g.  $(\Xi^x_m)^\prime x + (\Xi^u_m)^\prime u = \xi_m^\prime$. 
Since $J^\ast_0(x^+) = \frac{1}{2}\norm{x^+}_{P_0}^2$ is a given function, compatibility constraints are not necessary at 
stage $m=1$ and the maximisation problem is given by:
\begin{equation}
	\hat J_1^\ast(x,u) = \left\{\begin{split}
	\max_{w,x^+} & -\frac{\gamma^2}{2}\norm{w}^2 + J_{0}^\ast(x^+)\\
	\text{s.t.} & 
	\underbrace{\left(\begin{array}{cc}
	G^\prime & 0 \\ -D & I
	\end{array}\right)}_{\bar H}
	\left(\begin{array}{c}w\\ x^+\end{array}\right)
	\\&= \underbrace{\left(\begin{array}{cc}W_x^\prime & W_u^\prime \\ A & B\end{array}\right)}_{
	\bar S
	}
	\left(\begin{array}{c}x \\ u \end{array}\right) + \underbrace{\left(\begin{array}{c}
	w_c^\prime \\ 0\end{array}\right)}_{\bar f}
	\end{split}\right.
\end{equation}
with $W_x^\prime, W_u^\prime$ and $w_c^\prime$ denoting the active rows of the right hand side 
of~\eqref{eq:dist:cons}.
With Lemma~\ref{thm:mpQP:solution} we obtain
\begin{equation}
      \left(\begin{array}{c}
      w^\ast \\
      x^{+,\ast} \\
      \eta^\ast \\
      \lambda^\ast
      \end{array}\right) = \left(\begin{array}{c}M^w_z\\ M^{x^+}_z \\ N^{\eta}_z \\ N^{\lambda}_z
      \end{array}\right)\underbrace{\left(\begin{array}{c}x\\ u\end{array}\right)}_z + 
      \left(\begin{array}{c}0\\ 0 \\ N^{\eta}_{\hat\beta} \\ N^{\lambda}_{\hat\beta}
      \end{array}\right)\hat\beta + \left(\begin{array}{c}m^w\\ m^{x^+} \\ n^{\eta} \\ n^{\lambda}
      \end{array}\right)
\end{equation}
where $\eta^\ast$ denotes the dual variable to the active inequality constraints and hence must
be non-negative, i.e.~$\eta^\ast\geq0$. Notice that we omitted stage indices to avoid overloaded notation.
The cost-to-go for the minimisation problem at stage $m=1$ is
\begin{equation}
	\hat J_1^\ast(z) = \frac{1}{2}z^T
    \hat Q_1 z + c_1 z + d_1
\end{equation}
and the compatibility constraint is given by
\begin{equation}\label{prototype:compatibility:conditions}
	N_{\hat\beta}^T(\bar S z + \bar f) = 0,
\end{equation}
i.e.\ $C_1^x$ and $C_1^u$ are given as a row wise normalised version of~\eqref{prototype:compatibility:conditions}.
Hence the minimisation problem at $m=1$ becomes
\begin{equation}
	J_1^\ast(x) = \left\{\begin{split}
	\min_{u} & \frac{1}{2}\left(\norm{x}^2_Q + \norm{u}_R^2\right) + \hat J_{1}^\ast(x,u)\\
    \text{s.t.} & 
    \underbrace{\left(\begin{array}{c}
    (\Xi_1^u)^\prime \\ C_1^u
    \end{array}\right)}_{\bar C}
    u
    = \underbrace{\left(\begin{array}{c}-(\Xi^x_1)^\prime \\ -C_1^x\end{array}\right)}_{
    \bar T
    }x + 
    \underbrace{\left(\begin{array}{c}\xi_1^\prime\\  {\bf{1}} \end{array}\right)}_{\bar e}
    \end{split}\right.
\end{equation}
and its solution is given by
\begin{equation}
	\left(\begin{array}{c}
	u^\ast \\
	\nu^\ast \\
	\hat\zeta^\ast
	\end{array}\right) = \left(\begin{array}{c}K^u_x\\ O^{\nu}_x \\ O^{\hat\zeta}_x
	\end{array}\right)x + 
	\left(\begin{array}{c}0\\ O^{\nu}_\beta \\ O^{\hat\zeta}_{\beta}
	\end{array}\right)\beta + \left(\begin{array}{c}k^u\\ o^{\nu} \\ o^{\hat\zeta}
	\end{array}\right).
\end{equation}
Here $\hat C_2$ is a normalised version of $O^T_\beta(\bar T x + \bar e) = 0$.
For completeness, the $m=2$ maximisation and its solution are:
\begin{equation}
	\hat J_2^\ast(x,u) = \left\{\begin{split}
	\max_{w,x^+} & -\frac{\gamma^2}{2}\norm{w}^2 + J_{1}^\ast(x^+)\\
	\text{s.t.} & 
	\underbrace{\left(\begin{array}{cc}
	G^\prime & 0 \\ -D & I \\
	0 & \hat C_2
	\end{array}\right)}_{\bar A}
	\left(\begin{array}{c}w\\ x^+\end{array}\right)
	\\&= \underbrace{\left(\begin{array}{cc}W_x^\prime & W_u^\prime \\ A & B \\ 0 & 0\end{array}\right)}_{
	\bar S
	}
	\left(\begin{array}{c}x \\ u \end{array}\right) + \underbrace{\left(\begin{array}{c}
	w_c^\prime \\ 0 \\ {\bf{1}}\end{array}\right)}_{\bar b}
	\end{split}\right.
\end{equation}
\begin{equation}
	\left(\begin{array}{c}
	w^\ast \\
	x^{+,\ast} \\
	\eta^\ast \\
	\lambda^\ast\\
	\zeta^\ast
	\end{array}\right) = \left(\begin{array}{c}M^w_z\\ M^{x^+}_z \\ N^{\eta}_z \\ N^{\lambda}_z \\ N^{\zeta}_z
	\end{array}\right)z + 
	\left(\begin{array}{c}0\\ 0 \\ N^{\eta}_{\hat\beta} \\ N^{\lambda}_{\hat\beta} \\ N^\zeta_{\hat\beta}
	\end{array}\right)\hat\beta + \left(\begin{array}{c}m^w\\ m^{x^+} \\ n^{\eta} \\ n^{\lambda}\\ n^\zeta
	\end{array}\right)
\end{equation}
All subsequent minimisations and maximisations can be solved analogously provided the active set is
known.

\section{Exact solution of min-max program}\label{sec:line:search}
In the previous section we presented a recursive algorithm to solve~\eqref{eq:the:problem}
when all active constraints are known. In this section we present a method to find the active
constraints for a given initial state $\mathpzc x_e$ using a line search starting from a state $\mathpzc x_0$
for which the active constraints are known. In section~\ref{sec:gen:mpQP} and~\ref{sec:applied:mpQP}
we derived an explicit solution for each sub-optimisation program which depended affinely on the
state~$x$ at that stage~$m$ or the state and the input at that stage. Henceforth we denote the stage 
with an index, e.g.~$x_m$ denotes the optimised state in stage $m$, that is $x^{+,\ast}$ at stage $m+1$.
To ensure continuity of the dual variables we choose $\hat\zeta_m = \hat\beta_m$ and $\zeta_m=\beta_{m-1}$.
The affine character of the solution of each sub-problem allows us to parametrise the entire trajectory
with respect to parameters at stage~$m=N$, that is $x_N=\mathpzc x_0$ and $\beta_N$.

We recursively define all primal and dual variables:
\begin{equation}\label{eq:trajectory:affine:representation}
	\begin{split}
		u_N(x_N) =& \underbrace{K_{x,N}^x}_{\Psi_N^u} x_N + \underbrace{k_N^u}_{\psi_N^u}\\
		\nu_N(x_N,\beta_N) &= \underbrace{O_{x,N}^x}_{\Psi_N^\nu} x_N + \underbrace{O_{\beta,N}^\nu}_{\Phi_N^\nu} \beta_N + 
		\underbrace{o_N^\nu}_{\psi_N^u}\\
		&\vdots \\
		x_0(x_N) =& \underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right)}_{\Psi_1^x}x_N + 
		M_{z,1}^{x^+}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + m_1^{x^+}\\
		\eta_1(x_N,\beta_N) =& \underbrace{\left(N_{z,1}^{\eta}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \Psi_1^{\hat\zeta}\right)}_{\Psi_1^\eta}x_N + \underbrace{N_{\hat\beta,1}^\eta \Phi_1^{\hat\zeta}}_{\Phi_1^\eta} \beta_N\\
		&+\underbrace{N_{z,1}^{\eta}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \psi_1^{\hat\zeta} + n_1^\eta}_{\psi_1^\eta}\\
		&\vdots
	\end{split}
\end{equation}
Notice that $\Phi^{(\cdot)}_m$ is only non-trivial if the associated sub-problem is degenerate. The 
representation~\eqref{eq:trajectory:affine:representation} is valid as long as the constraint set remains unchanged, 
i.e. the inequalities that hold with equality do not change. Starting from the known set of active constraints at
$\mathpzc x_0$ we perform a line search towards $\mathpzc x_e$, i.e. we vary $x_N(\alpha)=\mathpzc x_0+
\alpha(\mathpzc x_e-x_0)$ for $0\leq\alpha\leq1$. We use~\eqref{eq:trajectory:affine:representation} to
check all inactive constraints with the primal variables $x_m(x_N(\alpha)),\, u_m(x_N(\alpha))$ and~$w_m(x_N(\alpha))$,
and all dual variables corresponding to inequality constraints to check whether constraints need to be deactivated. That is,
we search for the smallest $\alpha$ such that $\Xi_{m,i}^x x_m(x_N(\alpha))+\Xi_{m,i}^u u_m(x_N(\alpha))=\xi_{m,i}$
et cetera for the primal variables and the smallest $\eta_{m,i}(x_N(\alpha),\beta_N)=0$ and update the set of active
constraints accordingly, i.e. if an additional constraint is activated it is introduced to the equality constraint
mpQP recursion and the updated problems are solved again recursively, if a constraint needs to be deactivated
because the smallest $\alpha$ corresponds to the zero-crossing of its dual variable then again the mpQP recursion 
is solved again. The updated active set is the new known active set at $\mathpzc x_0 = x_N(\alpha)$ and the line-search
continues until the state~$\mathpzc x_e$ is in the same active set as $\mathpzc x_N(\alpha)$ where the line-search terminates.
To initialise we start at the origin $\mathpzc x_0=0$, since we know that no constraints are active. Notice that the right hand side 
of~\eqref{eq:dist:cons} is updated the same way, a line-search is used to determine the active affine representation.

If there are compatibility constraints on the initial state, i.e.\ if $\beta_N$ is non-trivial, then $\alpha$ cannot continue 
to  increase since equality constraints on $x_N(\alpha)$ are active, however $\beta_N$ can be used to update the active set. 
As in~\cite{Buerger:ACC} the free degeneracy variable $\beta_N$ is chosen to deactivate an active constraint on $x_N(\alpha)$ 
as long as no inequality constraints would be violated, i.e. as long as no dual variable becomes negative.

\section{Example}\label{sec:example}
This section we present a numerical simulation in which the state-dependent uncertainty enters through an upper bound on the linearisation
error for a nonlinear system. We consider an inverted pendulum described by $ml^2\ddot\varphi = mlg\sin\varphi+M$, where
$m=0.1$\,kg is the mass, $l=0.3$\,m is the length of the pendulum, $g=9.81$\,m\,s$\mbox{}^{-2}$ is the gravitational
constant and $M$ is the direct torque input. Discretisation using Euler backward finite differences 
at $f = 20$\,Hz yields
\begin{equation}
  \begin{split}
    x_1[k+1] &= x_1[k] + \frac{1}{f} x_2[k]\\
    x_2[k+1] &= \frac{g}{fl} \sin(x_1[k]) + x_2[k] + \frac{1}{fml^2}u[k]
  \end{split}
\end{equation}
where $x_1[k] = \varphi(t_0+k/f),x_2[k] = \dot\varphi(t_0+k/f)$ and $u[k] = M(t_0+k/f)$.
The nonlinearity is given by the sine function, and we can explicitly state the 
linearisation error as $e(x_1)=\frac{g}{fl}(\sin x_1 - x_1)$. Since the linearisation
only approximates the nonlinear dynamics around the equilibrium we can choose a closed interval 
$0\in D\subset\mathbb R$ and maximise the upper bound 
\begin{equation}
  \Delta e = \max_{x_1\in D}\abs{\frac{de}{dx_1}} = \frac{g}{fl}\max_{x_1\in D}\abs{\cos x_1-1}.
\end{equation}
This results in the linear system
\begin{equation}
  x^+ = \underbrace{\left(\begin{array}{cc}
  1 & f^{-1}\\ g(fl)^{-1} & 1
  \end{array}\right)}_A x + \underbrace{\left(\begin{array}{c} 0 \\(ml^2f)^{-1} \end{array}\right)}_B u
  +\left(\begin{array}{c}w_1\\ w_2\end{array}\right)
\end{equation}
with $\abs{w_1}\leq w_{1,\max}$ and $\abs{w_2}\leq\max\{w_{2,\max},\Delta e \abs{x_1}\}$.
We also introduce the input constraint $\abs{u}\leq3$\,N\,m and use the interval $D = [-\frac{3\pi}{4},
\frac{3\pi}{4}]$ and a horizon of $N=7$ to obtain the sequence of state constraints $\mathcal X_m$
depicted in Figure~\ref{fig:numerical:example}. The line search starting at the origin and
exploring towards $\mathpzc x_e=(-6,50)$ produces the optimal trajectories at the active-set switches
which are illustrated in blue. In pink a warm-started line search terminates on the boundary of the
feasible set before reaching the (infeasible) state $\mathpzc x_e=(0,30)$.

\begin{figure}
\includegraphics[width=.48\textwidth]{res}
\caption{The light blue sets depict the projected state constraint iterates~$\mathcal X_m$,
    whereas the blue and pink lines are the state trajectories produced along the two line
    searches. The pink line search eventually violates the state constraints and terminates.}
\label{fig:numerical:example}
\end{figure}
\section{Conclusion}\label{sec:conclusion}

\bibliographystyle{ifacconf}        % Include this if you use bibtex 
\bibliography{MyLib}       % and a bib file to produce the 
								                 % bibliography (preferred). The
                                 % correct style is generated by
                                 % Elsevier at the time of printing.

\end{document}